### 2019-04-25 Final/Project-TEMPLATE.tex
Adding paper to the repo. Easier to edit remotely.

% 
% Example file for conference proceedings in Springer Verlag's LNCS format
%
% Christian Jacob, September 2007
% (with LaTeX code provided by Navneet Bhalla)
%

\documentclass[runningheads]{llncs}
\usepackage{llncsdoc}
\usepackage{makeidx}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Preamble: Packages required for the paper
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{epstopdf}
\makeindex
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{document}

\title{Title of Your Project}
%\subtitle{<subtitle of your contribution>}
%\titlerunning{<Your abbreviated contribution title>} 
%\toctitle{<Your changed title for the table of contents>}

\author{Your Name\inst{1}  }
%\authorrunning{<abbreviated author list>}
%\tocauthor{<enhanced author list for the table of contents>}

\institute{ 
Dept. of Computer Science, Faculty of Science, University of Calgary, 2500 University Drive N.W., Calgary, Alberta, Canada T2N 1N4\\
\email{YourName@ucalgary.ca}
}

\maketitle

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% ABSTRACT
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{abstract} 
Describe in one or two sentences what your project is about.
\end{abstract}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% MAIN TEXT
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\section{Introduction}
\label{sec:Introduction}
Give an introduction to why you have chosen this project. Why is it of interest to study what you propose to explore in your project?

References are added through a bibliography file (e.g., References.bib), so that one has easy access to referenced items \cite{Alon:1999aa}.

\section{Related Work}
If you are aware of related work or similar work that has been done before, you should describe and mention it here. Here is also the place to refer to any papers, articles, web sites, etc. where you got your inspiration for your project from.

\section{Project Details}
Describe your project in more detail here. If you need more sophisticated formatting, have a look at the \emph{TEMPLATE.tex} file that came with this folder.

\section{Software Tools}
Mention the software tools you are going to use to implement your project. Be specific, i.e., add proper references.

\section{Time line}
Give an approximate time line for the implementation of your project.

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% References: Bibliography
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\bibliographystyle{splncs}
\bibliography{References} % add bibliography file here

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Index (optional): collects items in text from \index{...} command
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
%\printindex

\end{document}