### 2019-04-25 Final/paper.tex
Finished description of the phenotypical structure. Also outlined the rest of the implemenetation section.

\documentclass[runningheads]{llncs}
\usepackage{llncsdoc}
\usepackage{makeidx}

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% Preamble: Packages required for the paper
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{epstopdf}
\makeindex
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{document}

\title{Evolving Motion Machines}

\author{Nathan Douglas\\10160679}

\institute{ 
\email{nathan.douglas@ucalgary.ca}
}

\maketitle

% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
% ABSTRACT
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% \begin{abstract} 
% Describe in one or two sentences what your project is about.
% \end{abstract}

\section{Introduction}
\label{sec:Introduction}
Emergent computing has been previously used in many areas before.
In particular, the iterative approach taken by both evolutionary and genetic systems has repeatedly proven its' ability in several domains.
The gradual approximation technique for a given goal function can create very interesting, efficient, and effective solutions.
Further, some implementations have been remarkably effective in modelling real world behaviour.
With each implementation and domain being so unique, there exists plenty of room for continued application of these techniques.

This research aims to expand on the existing pool of implementations using emergent approaches.
General approaches are designed to either recreate natural behaviours or search for new optimal policies.
This is done first by creating a proposed solution, or genotype, with some initial value selected for each variable part of the solution.
The genotype or genotypes are then implemented and become their resulting phenotypes, which are then evaluated.
Finally, the genotypes which created the most effective phenotypes are mutated slightly, and repeat the process of instantiation and evaluation.
This creates the iterative process of searching for better and better solutions to a given goal function.

However, most research surrounding emergent computing is more focused on behaviours.
The structure of the phenotype is generally static, while the actions and reasoning are given more careful attention.
This system looks to use evolutionary strategies in a different way, focussing more on the resulting structure of a solution.
To this end, a rather simple goal function is used: simply move as much as possible.
The more complex aspect is in the genotype's large influence over the structure of the agent it generates.

Agents in this system are created from small, atomic pieces that join together to create larger collective structures.
Atoms have different simple geometric shapes, different types of motion, and a variable direction and intensity of the motion.
While atoms can mutate this direction and intensity as part of their evolutionary loop, they can also mutate additional adjoining atoms.
These new pieces attach to their parent using a physical link, creating a large branching tree structure.
With each atom still moving independently, this creates large agents with many independently moving subtrees.

This high degree on independence between subtrees creates a very dynamic structure with many different potential solutions.
Ideally this variability allows for interesting results.
As with evolution in nature, this system intends to mimic the variation in structure of life on earth.
As is shown below, this goal is somewhat met but the system itself is still too premature to be a comparable analog to natural selection.

\section{Implementation}
With the goal of moving as far as possible in each generation, an appropriate environment is needed to evaluate in.
Further, within the environment there are a list of features that must be included to support the evaluation.
First and foremost is the ability to deserialize a given genotype into a phenotype, or agent, to be evaluated.
Also needed is the inverse serialization ability.
Then, agents need some physics simulator to operate within and source the evaluation of exactly how much distance has been covered.
To this end, the Unity Engine is used as it provides a simple object instantiation scheme and a built-in physics simulation engine.

\subsection{Phenotypical Agent Structure}
Unity's simple object instantiation scheme is used to create the agent's constituent atoms.
Unity contains simple geometric shapes including spheres, cubes, and capsules.
Atoms can be one of these three shapes, which are scaled individually to make all atoms approximately the same size.
The scaling factor is different for each shape, as the shapes are different sizes by default.
The factor is set such that each shape has no cross section with a width larger than 1 meter as defined by Unity's measurement system.
This way, adjacent atoms can be instantiated 1 meter apart and do not overlap, as overlapping objects cause unexpected behaviour in the physics simulation.

Agents themselves are collections of these atoms.
Specifically, an agent is a 3-dimensional cubic grid, where each item in the grid can be either an atom or simply empty.
Atoms adjacent to each other within the grid can be physically linked to each other with invisible joints. 
These joints force the two connected atoms to be a static distance apart, mostly restriciting them from moving either closer or further apart.
The distance between two joined atoms is not completely static though, as Unity's physics engine allows the joints to stretch, compress, and bend to a small degree.

Each individual atom may either remain static or rotate around the center of its space in the grid.
This is the main way that an agent is able to move, outside of gravity causing an unbalanced agent to topple.
Further, the joint based structure combined with individual atom rotations allows for complex agent movement.
A rotating atom attached to another atom causes the attached atom to leave it's initial space in the grid and orbit the rotating atom.
Intuitively, this works like a fan.
A center piece rotates in place, and the blades are forced in a cyclic motion around the center.
The direction and force of rotation varies between atoms, with direction also controlling orientation for both static and non-static agents.

\subsection{Genotypical Agent Structure}

The genotypical structure of agents is very similar to the phenotypical instantiation.
Agents are stored as 3-dimensional arrays with each index contain specific data about its corresponding atom.

\subsection{Mutation System}

\subsection{Evaluation System}

\section{Results}

\section{Discussion}

\section{Future Work}

%different distance measurement (relative angle)
%different terrain types
%different evaluation point (center of mass)
\section{Conclusion}

\end{document}